---
title: "Development of a methodolygy to quantify compounds with Seal AA3"
date: "14 & 15 december 2017"
author: "Antoine Batigny & Guyliann Engels"

output: 
  html_notebook: 
    highlight: pygments
    theme: readable
    toc: yes
---



# Aim 

* We use Stareso sea water to prepare calibration line because we think that the SSW is contaminated


# Introduction

we use the Seal AA3 to compare several water's solutions (fresh water and sea water) with the addition of nitrogen and phosphorus compounds. 




We analyse :


- The Stareso sea water.


This water is aged with several cycles (light/ dark cycles). The organisms in water will use all available nutrients. this sea water is logically clean.

- The ultrapure water ( type 1, Millipore Corporation)


This water is described as ultrapure water (Type1). attention, the salinity is different between sample. a fresh water and sea water are not comparable.

- The Synthetic sea Water (SSW)


- The Stareso sea water with addition of 0.05, 0.1, 0.2, 1, 2, 5 and 10 $\mu mol/L$ for $PO_4$, $NO_3$ and $NH_4$ (method A). The Stareso sea water with addition of 0.05, 1, 30 $\mu mol/L$ for $Ptot$ and $Ntot$ (method B). The Stareso sea water with addition of 0.05, 1, 10 $\mu mol/L$ for $NO_2$. (method B)


# M&M

- Calibration lines are realised with Stareso sea water

- Solvant for seal AA3 are realsied with Stareso sea water

- each sample is measured 10 times (except calibration line)


# Results

```{r}
library(tidyverse)#version 1.1.1
library(ggpubr)#version0.1.6
library(gridExtra)#version 2.2.1
library(knitr)#version 1.17

data <- read.table(file = "../data/171215A.txt", sep = ";", header = TRUE) # import dataset

# check the method, line 8 , two possibility : method A (PO4, NO3, NH4) or method B (Ptot, Ntot, NO2)

# skip first line
meth_a <- data[ -c(1:13), ]
remove(data)
# rename variable for more simplicity
meth_a %>%rename(echant = ANAL, type  = X.1, date_analyse = X.5, 
                PO4_theo = X.6, PO4_conc = X.7, PO4_val = X.8,  
                NO3_theo = X.9, NO3_conc = X.10, NO3_val = X.11,
                NH4_theo = X.12, NH4_conc = X.13, NH4_val = X.14)%>%
  select(c(1,4,8:17))-> meth_a

# special attention for the type of variables and change if necessary
meth_a %>% mutate(echant = as.character(echant),
                  type = as.character(type),
                  PO4_theo = as.numeric(as.character(PO4_theo)),
                  PO4_conc = as.numeric(as.character(PO4_conc)),
                  PO4_val = as.numeric(as.character(PO4_val)),
                  NO3_theo = as.numeric(as.character(NO3_theo)),
                  NO3_conc = as.numeric(as.character(NO3_conc)),
                  NO3_val = as.numeric(as.character(NO3_val)),
                  NH4_theo = as.numeric(as.character(NH4_theo)),
                  NH4_conc = as.numeric(as.character(NH4_conc)),
                  NH4_val = as.numeric(as.character(NH4_val))) -> meth_a

meth_a$date_analyse <- as.POSIXlt(meth_a$date_analyse, format = "%d/%m/%Y %H:%M:%S")
meth_a$date_analyse <- as.POSIXct(meth_a$date_analyse)



# Import second dataset and check method

meth<- read.table(file = "../data/171214A.txt", sep = ";", header = TRUE) 

# check the method, line 8 , two possibility : method A (PO4, NO3, NH4) or method B (Ptot, Ntot, NO2)

## If METHOD B

# skip first line
meth_b <- meth[ -c(1:13), ]
remove(meth)
# rename variable for more simplicity

meth_b %>%rename(echant = ANAL, type  = X.1, date_analyse = X.5, 
                Ptot_theo = X.6, Ptot_conc = X.7, Ptot_val = X.8,  
                Ntot_theo = X.9, Ntot_conc = X.10, Ntot_val = X.11,
                NO2_theo = X.12, NO2_conc = X.13, NO2_val = X.14)%>%
  select(c(1,4,8:17))-> meth_b

# special attention for the type of variables and change if necessary
meth_b %>% mutate(echant = as.character(echant),
                  type = as.character(type),
                  Ptot_theo = as.numeric(as.character(Ptot_theo)),
                  Ptot_conc = as.numeric(as.character(Ptot_conc)),
                  Ptot_val = as.numeric(as.character(Ptot_val)),
                  Ntot_theo = as.numeric(as.character(Ntot_theo)),
                  Ntot_conc = as.numeric(as.character(Ntot_conc)),
                  Ntot_val = as.numeric(as.character(Ntot_val)),
                  NO2_theo = as.numeric(as.character(NO2_theo)),
                  NO2_conc = as.numeric(as.character(NO2_conc)),
                  NO2_val = as.numeric(as.character(NO2_val))) -> meth_b

meth_b$date_analyse <- as.POSIXlt(meth_b$date_analyse, format = "%d/%m/%Y %H:%M:%S")
meth_b$date_analyse <- as.POSIXct(meth_b$date_analyse)
```

## Calibration curves

```{r}
# select CALB to check a good calibration line
meth_a%>%filter(type == "CALB") -> calb_a
# If ok next step, otherwise compute a new variable with best calibration line between theorical values and obtained values

#PO4
PO4_graph <- ggplot(data = calb_a, mapping = aes(x = PO4_theo, y = PO4_val))+
  geom_point()+
  geom_smooth(method = "lm") + 
  labs( x = "Theoretical concentrations", y = "Values", caption = "PO4") +
  theme(plot.caption = element_text(hjust = 0.5)) +
  theme_bw()
# simply graph to visualize the calibration line for PO4

# Compute linear regression and extract r^2^ for PO4
PO4_reg <-lm(PO4_val ~ PO4_theo, data =  calb_a)

summary((PO4_reg))$coefficients -> PO4_tab
PO4_tab <- as.data.frame(PO4_tab)

round(summary((PO4_reg))$r.squared, 4) -> r_squared
PO4_tab$r_squared <- c(r_squared, " ")

# Same procedure for all elements
#NO3

NO3_graph <- ggplot(data = calb_a, mapping = aes(x = NO3_theo, y = NO3_val))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  labs( x = "Theoretical concentrations", y = "Values", caption = "NO3")+
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()

NO3_reg <-lm(NO3_val ~ NO3_theo, data =  calb_a)

summary((NO3_reg))$coefficients -> NO3_tab
NO3_tab <- as.data.frame(NO3_tab)

round(summary((NO3_reg))$r.squared, 4) -> r_squared
NO3_tab$r_squared <- c(r_squared, " ")

#NH4

NH4_graph <- ggplot(data = calb_a, mapping = aes(x = NH4_theo, y = NH4_val))+
  geom_point()+
  geom_smooth(method = "lm") + 
  labs( x = "Theoretical concentrations", y = "Values", caption = "NH4")+
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()


NH4_reg <-lm(NH4_val ~ NH4_theo, data =  calb_a)

summary((NH4_reg))$coefficients -> NH4_tab
NH4_tab <- as.data.frame(NH4_tab)

round(summary((NH4_reg))$r.squared, 4) -> r_squared
NH4_tab$r_squared <- c(r_squared, " ")


calb_tab <- bind_rows( PO4_tab, NO3_tab, NH4_tab)
calb_tab_a <- data_frame(coef= c( "I", "Val",  "I", "Val",  "I", "Val"))
calb_tab <- bind_cols(calb_tab_a, calb_tab)

## Template for tab 
mytheme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 0.6)),
    colhead = list(fg_params=list(cex = 0.7)),
    rowhead = list(fg_params=list(cex = 0.7)))
## Create tab with lm's coefficients 
tab <- tableGrob(calb_tab[ , - c(4)], theme = mytheme, rows = c("PO4", " ", "NO3", " ", "NH4", ""))

#multiplot
ggarrange(PO4_graph, NO3_graph, NH4_graph, tab)

remove(calb_tab, calb_tab_a, mytheme, NH4_graph, NH4_reg, NH4_tab, NO3_graph, NO3_reg, NO3_tab, PO4_graph, PO4_reg, PO4_tab, tab)
```

The r^2^ of calibration line are close  to 1. 

```{r}

# select CALB to check a good calibration line
meth_b%>%filter(type == "CALB") -> calb_b


#Ptot
Ptot_graph <- ggplot(data = calb_b[c(1:8), ], mapping = aes(x = Ptot_theo, y = Ptot_val))+
  geom_point()+
  geom_smooth(method = "lm") + labs( x = "Theoretical concentrations", y = "Values", caption = "Ptot")+
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()

Ptot_reg <-lm(Ptot_val ~ Ptot_theo, data =  calb_b[c(1:8), ])

summary((Ptot_reg))$coefficients -> Ptot_tab
Ptot_tab <- as.data.frame(Ptot_tab)

round(summary((Ptot_reg))$r.squared, 4) -> r_squared
Ptot_tab$r_squared <- c(r_squared, " ")

#Ntot
Ntot_graph <- ggplot(data = calb_b[c(1:8), ], mapping = aes(x = Ntot_theo, y = Ntot_val))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  labs( x = "Theoretical concentrations", y = "Values", caption = "Ntot")+
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()

Ntot_reg <-lm(Ntot_val ~ Ntot_theo, data =  calb_b[c(1:8), ])

summary((Ntot_reg))$coefficients -> Ntot_tab
Ntot_tab <- as.data.frame(Ntot_tab)

round(summary((Ntot_reg))$r.squared, 4) -> r_squared
Ntot_tab$r_squared <- c(r_squared, " ")

#NO2

NO2_graph <- ggplot(data = calb_b[c(9:17), ], mapping = aes(x = NO2_theo, y = NO2_val))+
  geom_point()+
  geom_smooth(method = "lm") + 
  labs( x = "Theoretical concentrations", y = "Values", caption = "NO2")+
  theme(plot.caption = element_text(hjust = 0.5))+
  theme_bw()


NO2_reg <-lm(NO2_val ~ NO2_theo, data =  calb_b[c(9:17), ])

summary((NO2_reg))$coefficients -> NO2_tab
NO2_tab <- as.data.frame(NO2_tab)

round(summary((NO2_reg))$r.squared, 4) -> r_squared
NO2_tab$r_squared <- c(r_squared, " ")


calb_tab <- bind_rows( Ptot_tab, Ntot_tab, NO2_tab)
calb_tab_b <- data_frame(coef= c( "I", "Val",  "I", "Val",  "I", "Val"))
calb_tab <- bind_cols(calb_tab_b, calb_tab)

## Template for tab 
mytheme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 0.6)),
    colhead = list(fg_params=list(cex = 0.7)),
    rowhead = list(fg_params=list(cex = 0.7)))
## Create tab with lm's coefficients 
tab <- tableGrob(calb_tab[ , - c(4)], theme = mytheme, rows = c("Ptot", " ", "Ntot", " ", "NO2", ""))

#multiplot
ggarrange(Ptot_graph, Ntot_graph, NO2_graph, tab)

remove(calb_tab, calb_tab_b, mytheme, NO2_graph, NO2_reg, NO2_tab, Ntot_graph, Ntot_reg, Ntot_tab, Ptot_graph, Ptot_reg, Ptot_tab, tab)
```

The r^2^ of calibration line round to 1. 


## Samples 

```{r}
#meth A 
meth_a%>%filter(type == "SAMP") -> samp_a # select only samp (so no calibration or drift points)


samp_a %>%
  filter(echant != "Milliq") -> samp_a

samp_a$echant <- factor(x = samp_a$echant,ordered = TRUE,levels = c("H2O Stareso", "SSW", "cal0.05", "cal0.1", "cal0.2", "cal1", "cal2", "cal5", "Cal10")) #arrange the echant by concentrations


#meth B

meth_b%>%filter(type == "SAMP") -> samp_b

## Recodage de samp_b$echant
samp_b$echant <- as.character(samp_b$echant)
samp_b$echant[samp_b$echant == "Milliq_____________"] <- "Milliq"
samp_b$echant[samp_b$echant == "ssw"] <- "SSW"

samp_b %>%
  filter(echant != "Milliq") -> samp_b

samp_b$echant <- factor(x = samp_b$echant, ordered = TRUE, levels = c("H2O Stareso", "Milliq", "SSW","cal0.05 N+P", "cal1 N+P", "cal30 N+P" ,"cal0.05 NO2", "cal1 NO2", "cal10 NO2" )) #arrange the echant by concentrations


```


### PO4

It is possible to work with values or values that is converted by calibration line. We first work with concentrations.
As you can see on the graph, the aged sea water is less concentrated that Milli-Q (ultrapure water of type 1). Furthermore, the SSW (synthetic sea water) is more concentrated that Milli-Q.

```{r}
ggplot(data = samp_a, mapping = aes(x = echant,y = PO4_val))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Values", caption = "Phosphate's values of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))
ggplot(data = samp_a, mapping = aes(x = echant,y = PO4_conc))+
  geom_boxplot()+ 
  labs( x = "Samples", y = "Concentration (µmol/L)", caption = "Phosphate's concentration (µmol/L) of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))
```

The boxplots show several extreme points. We remove theses points. 

```{r}
red <- samp_a[ -c(4, 19:20) , ]

red %>% 
  group_by(echant)%>%
  summarise(mean = round(mean(PO4_conc), digits = 4), 
            sd = round(sd(PO4_conc), digits = 4),cv = round(x = ((sd/mean)*100), digits = 4) ,  
            number = length(PO4_conc)) -> red

kable(x = red , format = "pandoc", caption = "Mean, standard deviation (sd) and variation coefficient (cv) of phosphate's concentrations of different samples.")
```

In the calibration line, the values of Stareso sea water is of about 2800 ($0 \mu mol/L$) whereas the values of Stareso sea water in sample is of about 600 ($-0.42\mu mol/L$). 


### NO3

```{r}
ggplot(data = samp_a, mapping = aes(x = echant,y = NO3_val))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Values", caption = "Nitrate's values of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))


ggplot(data = samp_a, mapping = aes(x = echant,y = NO3_conc))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Concentration (µmol/L)", caption = "Nitrate's concentration (µmol/L) of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))
```



```{r}
samp_a %>% 
  group_by(echant)%>%
  summarise(mean = round(mean(NO3_conc), digits = 4), 
            sd = round(sd(NO3_conc), digits = 4),cv = round(x = ((sd/mean)*100), digits = 4) ,  
            number = length(NO3_conc)) -> a

kable(a, format = "pandoc", caption = "Mean, standard deviation (sd) and variation coefficient (cv) of nitrate's concentrations of different samples.")
```

In the calibration line, the values of Stareso sea water is of about 3500 ($0 \mu mol/L$) whereas the values of Stareso sea water in sample is of about 3600 ($0.068\mu mol/L$). 
The table shows that above  $0.05 \mu mol/L$ the coefficient variation is lower that 5%.

### NH4

```{r}
ggplot(data = samp_a, mapping = aes(x = echant,y = NH4_val))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Values", caption = "Ammonium's values of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))

ggplot(data = samp_a, mapping = aes(x = echant,y = NH4_conc))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Concentration (µmol/L)", caption = "Ammonium's concentration (µmol/L) of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))
```


Several values are zero values except SSW, 2 , 5 and 10 $\mu mol/L$.



### Ptot

```{r}
ggplot(data = filter(samp_b, echant == c("H2O Stareso", "SSW","cal0.05 N+P", "cal1 N+P","cal0.05 NO2")), mapping = aes(x = echant,y = Ptot_val))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Values", caption = "Phosphorus's values of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))

ggplot(data = filter(samp_b, echant == c("H2O Stareso", "SSW","cal0.05 N+P", "cal1 N+P", "cal0.05 NO2")), mapping = aes(x = echant,y = Ptot_conc))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Concentration (µmol/L)", caption = "Phosphorus's concentration (µmol/L) of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))
```




```{r}

red <- samp_b[ -c(6, 10) , ]

red %>% 
  group_by(echant)%>%
  summarise(mean = round(mean(Ptot_conc), digits = 4), 
            sd = round(sd(Ptot_conc), digits = 4),cv = round(x = ((sd/mean)*100), digits = 4) ,  
            number = length(Ptot_conc)) -> a

kable(a, format = "pandoc", caption = "Mean, standard deviation (sd) and variation coefficient (cv) of phosphorus' concentrations of different samples.")
```



### Ntot

```{r}

ggplot(data = filter(samp_b, echant == c("H2O Stareso",  "SSW","cal0.05 N+P", "cal1 N+P","cal0.05 NO2")), mapping = aes(x = echant,y = Ntot_val))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Values", caption = "Nitrogen's values of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))


ggplot(data = filter(samp_b, echant == c("H2O Stareso",  "SSW","cal0.05 N+P", "cal1 N+P", "cal0.05 NO2")), mapping = aes(x = echant,y = Ntot_conc))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Concentration (µmol/L)", caption = "Nitrogen's concentration (µmol/L) of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))


```

```{r}
samp_b %>% 
  group_by(echant)%>%
  summarise(mean = round(mean(Ntot_conc), digits = 4), 
            sd = round(sd(Ntot_conc), digits = 4),cv = round(x = ((sd/mean)*100), digits = 4) ,  
            number = length(Ntot_conc)) -> a

kable(a, format = "pandoc", caption = "Mean, standard deviation (sd) and variation coefficient (cv) of Nitrogen's concentrations of different samples.")
```


### NO2

```{r}
ggplot(data = samp_b, mapping = aes(x = echant,y = NO2_val))+
  geom_boxplot()+
  geom_boxplot() + 
  labs( x = "Samples", y = "Values", caption = "Nitrite's values of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))

ggplot(data = samp_b, mapping = aes(x = echant,y = NO2_conc))+
  geom_boxplot() + 
  labs( x = "Samples", y = "Concentration (µmol/L)", caption = "Nitrite's concentration (µmol/L) of different samples")+
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0.5))
```


```{r}
samp_b %>% 
  group_by(echant)%>%
  summarise(mean = round(mean(NO2_conc), digits = 4), 
            sd = round(sd(NO2_conc), digits = 4),cv = round(x = ((sd/mean)*100), digits = 4) ,  
            number = length(NO2_conc)) -> a

kable(a, format = "pandoc", caption = "Mean, standard deviation (sd) and variation coefficient (cv) of Nitrite's concentrations of different samples.")
```


# Discussion & perspective

The SSW (water that uses to prepare calibration line and blank) must be purified. 


The sea water is obtained by the mixing of Milli-Q, [sodium chloride](https://uk.vwr.com/store/product/777308/vwrc27810) ( $NaCl$, AnalaR Normapur ACS) and  [sodium hydrogen carbonate](https://uk.vwr.com/store/product/776026/vwrc27778) ($NaHCO_{3}$, AnalaR Normapur ACS). $35g/L$ of $NaCl$ and $0.2 g/L$ of $NaHCO_{3}$ are added to Milli-Q. 
  
  The sodium chloride and sodium hydrogen carbonate are not clean. There are a slight contamination. In $NaCl$, you can find 5ppm  of phosphate  total Nitrogen at most. You can also find  30ppm of a mixing between $ClO_3$ and $NO_3$. The contaminants of $NaHCO_{3}$ are less important.
    
* For 5ppm of $PO_4$ : 

$$5mg \ PO4 \Longrightarrow \ 1kg \ NaCl$$
$$35g \ NaCl + \ 1L \ MilliQ = synthetic \ sea \ water $$
$$175\mu g \Longrightarrow \ 1L\  synthetic \ sea \ water$$
$$\frac{175\mu g/L}{94 \ g/mol} = 1.86 \ \mu mol/L$$
The same equations can be using for total nitrogen. **Attention, the molar mass changes**. 

$$\frac{175 \ \mu g/L}{14 \ g/mol} = 12.5 \ \mu mol/L$$


In the SSW, you can fin $1.86 \ \mu mol/L$ of $PO_4$ and $12.5 \ \mu mol/L$ of total nitrogen at most. It is impossible to know the importance of contaminants.
    



IT is important that find a method to age SSW. 

The new experiment must be prepared to check the difference between 0 (in calibration line) and 0 in Stareso sea water samples. 


